---
title: "**AdvStDaAn, Worksheet, Week 7**"
author: "Michael Lappert"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: 3
  '': default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('Specific_R_functions/RFn_Plot-lmSim.R')
source('Specific_R_functions/RFn_Plot-glmSim.R')
```

## Exercise 1
### Exercise 1.a)
Turbine Data (cf. Exercise 1 on Worksheet Week 4) Does the GLM that you have ﬁtted in part 1(b) model the data adequately?
#### Dataset loading and sanity check:

```{r, out.width='100%'}
path <- file.path('Datasets', 'turbines.dat')
df <- read.table(path, header=TRUE)

# Fitted model in w4, 1.b)
glm1.1 <- glm(cbind(Fissures, Turbines-Fissures) ~ Hours, family = binomial, data = df)
summary(glm1.1)
```

Because the response is binomially distributed with m > 1, we can test on overdispersion:
```{r}
1-pchisq(10.331, 9) # if resulting value > 0.05 -> no overdispersion
```

Because the p-value is > than the significance level of 5% we have no evidence against the null hypothesis that $\phi$ = 1 -> no overdispersion.
```{r}
qchisq(0.95, df=9) # if resulting value > Residual deviance -> no overdispersion
```

Because the residual deviance is smaller than $q^{\chi^2_9}_{0.95}$ the null hypothesis that $\phi$ = 1 cannot be rejected -> no overdispersion.\

***
### Question 1.a)
*Is the conclusion in the two cells abvoe right? Are these two different methods to come to the same result (looking for overdispersion)?*

***

Performing a residual and sensitivity analysis for the fitted model:
```{r}
par(mfrow=c(2,4))
plot(glm1.1)
plot.glmSim(glm1.1, SEED = 1)
```

