---
title: "**AdvStDaAn, Worksheet, Week 2**"
author: "Micheal Lappert"
date: '05.04.2022'
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: 3
  '': default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('Specific_R_functions/RFn_Plot-lmSim.R')
```

## Exercise 1
#### Dataset loading and sanity check:

```{r, out.width='100%'}
path <- file.path('Datasets', 'Synthetic.dat')
df <- read.table(path, header=TRUE)

summary(df)
dim(df)
head(df)
tail(df)
plot(df)
```

There seems to be some strong correlation between x2 and Y but withing x1 and x2 seems not to be a problem with multicollinearity. We fit an robust MM-Estimator model to the data.
```{r}
library(robustbase)
rlm1.1 <- lmrob(Y ~ x1 + x2, data = df)
summary(rlm1.1)
coef(rlm1.1)
```

8 observations were identified as outliers from the MM-Estimator. The $R^{2}$ has a pretty good score of 0.986.

Coefficients:
```{r}
coef(rlm1.1)
```
 
The estimated standard dviation of the error is 1.111.

Residual ans sensitivity analysis:
```{r}
par(mfrow=c(2,4))
plot(rlm1.1)
plot.lmrob(rlm1.1)
```

