---
title: "**AdvStDaAn, Worksheet, Week 12**"
author: "Michael Lappert"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: 3
  '': default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales)
library(rmutil)
```

## Task 1
Approximate the integral\

$\int_{-1}^{2} \big(x^2*cos(0.6x)+0.2 \big)dx$\

via Monte Carlo simulation.

```{r}
# from solutionsheet:
f = function(x){
  return(x^2*cos(.6*x)+.2)
  }
x = seq(-1,2,.01)
x
y = f(seq(-1,2,.01))
y
plot(x, y, ylim=c(0,2), xlab="x", ylab="f(x)")
abline(h=c(0,2), v=c(-1,2))
n = 1000
hits = 0
for(i in 1:n)
{
    x_ = runif(1,-1,2)
    y_ = runif(1,0,2)
    col="black"
    if(y_<f(x_))
    {
        hits = hits+1
        col="green"
    }
    points(x_, y_, pch=16, col=col)
}
# hits/n = I/6  =>  I = hits/n*6
hits/n*6 # Exact Integral: 2.50
```

## Task 2
Proof: The stationary distribution is $(\frac{9}{27}, \frac{8}{27},  \frac{10}{27})$.\
Does this Markov model satisfy 'detailed balance'?

***

stationary distribution:\

9/27  = 9/27 * 0.0 + 8/27 * 0.5 + 10/27 * 0.5  ok \
8/27  = 9/27 * 0.8 + 8/27 * 0.1 + 10/27 * 0.0  ok \
10/27 = 9/27 * 0.2 + 8/27 * 0.4 + 10/27 * 0.5  ok \

$=> (9/27,8/27,10/27)$ is stationary distribution. \

detailed balance: \

consider state 2 and 3: 8/27 * 0.4 is not 10/27 * 0 \
$=>$ detailed balance does not hold.

## Task 3
Find the equilibrium distribution.

Does this Markov model satisfy ’detailed balance’?

***

From solution:

Stationary distribution:
```{r}
m = matrix(c(.5, 0, 1, .25, .5, .0, .25, .5, 0), 3)
m
m = m%*%m
m
m = m%*%m
m
# m doesn't change any more
# => (0.5,0.25,0.25) is stationary distribution
# proof this!
```

proof:\

0.50 = 0.50 * 0.5  + 0.25 * 0.0 + 0.25 * 1.0  ok \
0.25 = 0.50 * 0.25 + 0.25 * 0.5 + 0.25 * 0.0  ok \
0.25 = 0.50 * 0.25 + 0.25 * 0.5 + 0.25 * 0.0  ok \

detailed balance:  \

consider state 2 and 3: 0.25 * 0.5 is not 0.25 * 0 \
$=>$ detailed balance does not hold.

## Task 4
Find the equilibrium distribution.

Does this Markov model satisfy ’detailed balance’?

What’s the expected number of sunny days per year?

***

Stationary distribution:
```{r}
m = matrix(c(0.8, 0.4, 0.2, 0.6), 2)
m
m = m %*% m
m
(m = m %*% m)

#           [,1]      [,2]
# [1,] 0.6666667 0.3333333
# [2,] 0.6666667 0.3333333
# -> does not change anymore
# (0.67, 0.33) is the stationary distribution
```

Proof:
0.67 = 0.67*0.8 + 0.33*0.4 = `r 2/3*0.8 + 1/3*0.4` -> ok
0.33 = 0.67*0.2 + 0.33*0.6 = `r 2/3*0.2 + 1/3*0.6` -> ok

Detailed balance: \
2/3 * 0.2 = `r 2/3*0.2` = 1/3 * 0.4 = `r 1/3*0.4` -> satisfies detailed balance


The expected number of sunny days per year is 365 * 2/3 = `r 365 * 2/3`.

## Task 5
Recall our salmon selling company example: 6 out of 16 persons signed up.

Infer the sign-up rate with a Markov chain Monte Carlo method. Consider a uniform and a Beta(2, 20) prior on the sign-up-rate and compare your results with the exact results from the theoretical approach using a conjugate prior.

***

From solution:

**uniform prior:** \

```{r task5, eval = TRUE}
# 6 from 16 sign up for salomon
# Baysian Data Analysis via MCMC
nrSignups =  6
nrTotal   = 16

nSimulations = 100000

signUpRate_i = runif(1)
signUpRate_i

sample_signUpRates = rep(NA,nSimulations)

for(i in 1:nSimulations)
{
    signUpRate_prop = runif(1)
    u = runif(1)
    if( u < (dbinom(nrSignups,nrTotal,signUpRate_prop)/dbinom(nrSignups,nrTotal,signUpRate_i)))
    {
        signUpRate_i = signUpRate_prop
    }
    sample_signUpRates[i] = signUpRate_i
}


# chain with high correlation:
hist(sample_signUpRates,breaks=seq(0,1,.02),col="blue")
plot(sample_signUpRates)
plot(sample_signUpRates[1:100])

# Result:
signUpRates_iid = sample_signUpRates[seq(100,nSimulations,50)]
hist(signUpRates_iid,breaks=seq(0,1,.02),col="blue")
abline(v=nrSignups/nrTotal,col="yellow",lwd=3)
abline(v=mean(signUpRates_iid),col="red",lwd=3)
abline(v=quantile(signUpRates_iid,c(0.025,.975)),col="red",lty=3,lwd=3)

# Compare with theoretical result:
hist(signUpRates_iid,breaks=seq(0,1,.02),col="blue",freq=FALSE)
abline(v=nrSignups/nrTotal,col="yellow",lwd=3)
abline(v=mean(signUpRates_iid),col="red",lwd=3)
abline(v=quantile(signUpRates_iid,c(0.025,.975)),col="red",lty=3,lwd=3)
# theoretical results in green
curve(dbeta(x,1+6,1+10),add=T,lwd=3,col="green")
abline(v=7/18,col="green",lwd=1)
abline(v=qbeta(c(0.025,.975),7,11),col="green",lty=3,lwd=1)

```




**Beta(2, 20)-prior:** \

```{r task5b, eval = TRUE}
# 6 from 16 sign up for salomon
# Baysian Data Analysis via MCMC
nrSignups =  6
nrTotal   = 16

nSimulations = 100000

signUpRate_i = rbeta(1, 2, 20)

sample_signUpRates = rep(NA,nSimulations)

for(i in 1:nSimulations)
{
    signUpRate_prop = runif(1)
    u = runif(1)
    if( u < (dbinom(nrSignups,nrTotal,signUpRate_prop)*dbeta(signUpRate_prop,2,20))
             /(dbinom(nrSignups,nrTotal,signUpRate_i)*dbeta(signUpRate_i,2,20)) )
    {
        signUpRate_i = signUpRate_prop
    }
    sample_signUpRates[i] = signUpRate_i
}


# chain with high correlation:
#hist(sample_signUpRates,breaks=seq(0,1,.02),col="blue")
#plot(sample_signUpRates)
#plot(sample_signUpRates[1:100])

# Result:
signUpRates_iid = sample_signUpRates#= sample_signUpRates[seq(100,nSimulations,50)]
hist(signUpRates_iid,breaks=seq(0,1,.02),col="blue")
abline(v=nrSignups/nrTotal,col="yellow",lwd=3)
abline(v=mean(signUpRates_iid),col="red",lwd=3)
abline(v=quantile(signUpRates_iid,c(0.025,.975)),col="red",lty=3,lwd=3)

# Compare with theoretical result:
hist(signUpRates_iid,breaks=seq(0,1,.02),col="blue",freq=FALSE)
abline(v=nrSignups/nrTotal,col="yellow",lwd=3)
abline(v=mean(signUpRates_iid),col="red",lwd=3)
abline(v=quantile(signUpRates_iid,c(0.025,.975)),col="red",lty=3,lwd=3)
# theoretical results in green
curve(dbeta(x,2+6,20+10),add=T,lwd=3,col="green")
abline(v=8/38,col="green",lwd=1)
abline(v=qbeta(c(0.025,.975),8,30),col="green",lty=3,lwd=1)

```


## Task 6
Let 

x = c(3.4, -4.2, -0.7, -2.6, -1.6, -1.2, -2.2, -3.7, -0.9, -3.1) 

be some observations with corresponding responses

y = c(5.7, -0.1,  4.7,  2.7,  3.1,  2.3,  3.4,  1.0,  2.2,  1.5).

The values were simulated by the formular: y = 3 + 0.3*x + rnorm(10,0,1). \

Consider a linear model, i.e. $y_i = \beta_0 + \beta_1 x_i + \epsilon_i$, with Gaussian noise $\epsilon_i \sim Norm(o,\sigma^2)$. 
Assume a Gamma(1,1)- prior on sigma and normal distributed priors with mean 0 and standard deviation 3 on  $\beta_0$ and $\beta_1$.\

Perform a Bayesian analysis: Estimate the posterior mean and credible intervals for $\beta_0$, $\beta_1$ and sigma.

***

From solution:
```{r}
# Observed data:
obs_x = c(3.4, -4.2, -0.7, -2.6, -1.6, -1.2, -2.2, -3.7, -0.9, -3.1)
obs_y = c(5.7,-0.1, 4.7, 2.7, 3.1, 2.3, 3.4, 1.0, 2.2, 1.5)
# Log of (unnormalized) posterior density
logPosterior = function(b0_, b1_, sigma_){
   sum(dnorm(b0_+b1_*obs_x - obs_y, 0, sigma_, log=TRUE)) +
    dnorm(b0_, 0, 3, log=TRUE) +
    dnorm(b1_, 0, 3, log=TRUE) +
    dgamma(sigma_, 1, 1, log=TRUE)
}
# Starting value
b0 = runif(1,0,1)
b1 = runif(1,0,1)
sigma = runif(1,0,1)
# Collect sampled values
b0_sample = c()
b1_sample = c()
sigma_sample = c()
# Bayesian Data Analysis via MCMC
for(i in 1:10000) {
    b0_prop = rnorm(1, b0, 0.3)
    b1_prop = rnorm(1, b1, 0.3)
    s_prop = abs(rnorm(1, sigma, 0.3))
    R = exp(logPosterior(b0_prop, b1_prop, s_prop)-logPosterior(b0, b1, sigma))
    u = runif(1)
    if( u <  R)
    {
        b0 = b0_prop
        b1 = b1_prop
        sigma = s_prop
    }
    b0_sample = c(b0_sample,b0)
    b1_sample = c(b1_sample,b1)
    sigma_sample = c(sigma_sample,sigma)
}
# Evaluation
plot(b0_sample, main='Traceplot b0')
plot(b1_sample, main='Traceplot b1')
plot(sigma_sample, main='Traceplot sigma')
plot(b0_sample,b1_sample, main='Scatterplot b0 against b1')
# Thinning the chains to get iid samples
b0_iid = b0_sample[1:100*100]
b1_iid = b1_sample[1:100*100]
sigma_iid = sigma_sample[1:100*100]
plot(obs_x,obs_y)
for(i in 1:100)
{
  abline(b0_iid[i],b1_iid[i])
}
points(obs_x,obs_y,col='red')
mean(b0_iid)
quantile(b0_iid,c(0.025,.975))
mean(b1_iid)
quantile(b1_iid,c(0.025,.975))
mean(sigma_iid)
quantile(sigma_iid,c(0.025,.975))
# Compare
lm(obs_y~obs_x)
```




```{r}
# Code from slide 60 of week 13
x = -10 
s = 0.5 
xPath = c() 

for(i in 1:10000) {
  xProp = x + runif(1,-s,s)
  if(exp(-xProp^2/2)/exp(-x^2/2) > runif(1)) {
    x = xProp
    }
  xPath = c(xPath,x)
} 

plot(xPath) 
points(1:100*100, xPath[1:100*100], col="red") 
hist(xPath[1:100*100], breaks = 50)

runif(1)
```


## Task 2




***

### Question Task
|>Q1: 
|>>A1: 

***
