---
title: "**AdvStDaAn, Worksheet, Week 11**"
author: "Michael Lappert"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: 3
  '': default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales)
```

## Task 1
You can choose from two different types of light-emitting diodes.

You order 100 diodes of type 1 and 100 diodes of type 2. 8 diodes of type 1 and 12 diodes of type 2 were insufﬁcient for your purposes. However diodes of type 2 are 10% cheaper.

Is it beneﬁcial to buy diodes of type 2? Reject rates usually vary around 10% ± 5%. Propose an appropriate prior and perform a Bayesian analysis.

***
First one can calculate a and b for the beta prior.

$\frac{a}{a+b}$ = 0.1
b = 9a

$\frac{ab}{(a+b+1)(a+b)^{2}} = 0.05^{2}$

set b = 9a

$\frac{a*9a}{(a+9a+1)(a+9a)^{2}} = 0.05^{2}$ 
a = 3.5
b = 31.5

```{r}
# Simulate reject rates for diodes of type 1 and type 2
nSamples = 100000
postType1 = rbeta(nSamples, 3.5 +  8, 31.5 + 92)
postType2 = rbeta(nSamples, 3.5 + 12, 31.5 + 88)

hist(postType1, xlim = c(0, .5), col = 'blue', main = 'Histogram overlay of both priors')
hist(postType2, xlim = c(0, .5), col = alpha('red', 0.5), add = TRUE)

plot(postType1, postType2)
abline(0,1,col="blue")
sum(postType1< postType2)/nSamples

# Calculate expected reject rates
expType1 = (3.5 +   8)/ (3.5 +   8 +31.5 + 92) # -> E(X)
expType2 = (3.5 +  12)/ (3.5 +  12 +31.5 + 88) # -> E(X)

# To compare, include that diodes of type 2 are 10% cheaper
(1-expType1)
(1-expType2)/.9
```

=> Type 2 is better than Type 1.

## Task 2
***
## Question Task 1
|>Q: 


***